---
title: "Question 6: Does the number of the case change along with the area?"
format: 
    html:
        embed-resources: true
---
# Data Munging
```{r}
library(readxl)
library(caret)
library(MASS)
library(tidyverse)
```

```{r}
data <- read_excel('311_Cases_cleaned.xlsx')
data$neighborhood <- as.factor(data$neighborhood)
head(data)
```

```{r}
# Create a contingency table shows the count of each case type in each neighborhood
case_counts <- xtabs(~ case + neighborhood, data = data)
head(case_counts)
```

```{r}
case_counts_df <- as.data.frame(case_counts)
names(case_counts_df) <- c('case_type', 'neighborhood', 'count')
head(case_counts_df)
```


# Model
```{r}
# Now fit the Negative Binomial model
# The glm.nb function fits a negative binomial model which can be more appropriate for over-dispersed count data
model <- glm.nb(count ~ neighborhood + case_type, data = case_counts_df)
summary(model)
```

Positive coefficient like Mission has a coefficient of 1.71363, suggesting that being in the Mission neighborhood is associated with an increase in case counts compared to the baseline neighborhood
Negative coefficient like Seacliff has a coefficient of -2.15708, indicating fewer cases compared to the baseline neighborhood

Many of the neighborhoods and case types have significant p-values, indicating strong evidence against the null hypothesis.

Areas like Mission and South of Market show a positive impact on case counts, which could indicate higher activity or reporting rates in these areas. In contrast, areas like Seacliff and Presidio Heights show a negative impact, suggesting lower case counts.

Certain types of cases, such as Graffiti and Street and Sidewalk Cleaning, have high positive coefficients, indicating these are common complaints across the city. Conversely, case types like General Request - AGING ADULT SERVICES and General Request - CITY ATTORNEY have large negative coefficients, suggesting these are less frequently reported or specific to fewer areas.

```{r}
library(ggplot2)
library(ggmap)
```

```{r}
data_2012_cleaning <- subset(data, year == 2012 & case == "Street and Sidewalk Cleaning")
```

```{r}
case_counts <- table(data_2012_cleaning$neighborhood)

# Sorting the neighborhoods by the number of cases to make the histogram more informative
sorted_counts <- sort(case_counts, decreasing = TRUE)

barplot(sorted_counts, 
        main = "Histogram of Street and Sidewalk Cleaning Cases by Neighborhood (2012)",
        xlab = "Neighborhood",
        ylab = "Number of Cases",
        las = 2, # makes the labels perpendicular to axis
        col = "blue") 
```